Seasonality in dominant upper-ocean dynamics at submesoscales

### 2016-05-09 writeup/rocha_etal.tex
Changes fig2

@@

Focusing on the lateral gradients of the horizontal velocity vector.
The lateral gradients are calculated using a centered
second-order finite differences scheme. We then calculate
the vertical vorticity $\zeta \equiv v_x - u_y$,
horizontal divergence $\delta \equiv u_x + v_y$, and the magnitude
of the lateral rate of strain $S \equiv [(v_x+u_y)^2 + (u_x-v_y)^2]^{1/2}$.
These quantities highlight the submesoscale structures in the flow
and hint on their dynamics.

@@

The horizontal divergent part of the flow is significant: the RMS is $\delta/f > 0.1$).
Most of these horizontal divergent flows project are supra-inertial and project
on scales smaller than 100 km: daily-averaging the velocity fields or low-pass
filtering (100 km cutoff) reduces the horizontal divergence by about 70$\%$.
Furthermore, the seasonal cycle in horizontal divergence is out-of-phase with that
of vertical vorticity: the horizontal divergence is about 30$\%$ larger in late
Summer/early Fall than it is in late Winter/early Spring. The divergence associated
with submesoscale subinertial eddies is small and its seasonal cycle is in phase
with that of vorticity (see green lines).  The seasonal cycle of the horizontal
divergence is roughly in phase with the seasonal cycle of the near-surface (0-50m)
stratification.
\begin{figure*}[ht]
\begin{center}
\includegraphics[width=1.\textwidth]{figs/fig1.pdf}
 \caption{Snapshots of surface vorticity (a and b) and a transect
          of potential density at 165$^\circ$E (c and d). The snapshots were
          taken at 00:00 UTC.}
 \label{fig1}
 \end{center}
 \end{figure*}

 \begin{figure*}[ht]
   \begin{center}
     \includegraphics[width=.9\textwidth]{figs/fig2.pdf}
  \caption{Time series of the root-mean-square of kinetic energy (a), vorticity
  (b), lateral strain (c), horizontal divergence (d), and stratification in the upper 50 m (e).}
  \label{fig2}
  \end{center}
\end{figure*}

@@

\section{Projection on time scales}
To better understand the which flows project onto high-wavenumbers, the projection
of these flows on time scales, we calculate frequency spectra of surface KE
(figure \ref{fig4}). To obtain statistically meaningful results, we average
spectral estimates along a section at 165$^\circ$E.  There is also about 50
independent realizations of spectral estimates.

@@

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=.45\textwidth]{figs/fig4.pdf}
 \caption{Surface (horizontal) KE frequency spectra. Solid lines
 are spectra based on the 1/24$^\circ$ fields, dashed line are spectra
 based on 100-km-low-pass smoothed fields.}
 \label{fig4}
 \end{center}
\end{figure}

>>>>

% CR added this
\usepackage{color}

@@

Focusing on the lateral gradients of the horizontal velocity vector.
The lateral gradients are calculated using a centered
second-order finite differences scheme. We then calculate
the vertical vorticity $\zeta \equiv v_x - u_y$ and 
horizontal divergence $\delta \equiv u_x + v_y$.
These quantities highlight the submesoscale structures in the flow
and hint on their dynamics.

@@

The horizontal divergent part of the flow is significant: the RMS is $\delta/f > 0.1$).
Most of these horizontal divergent flows project are supra-inertial and project
on scales smaller than 100 km: daily-averaging the velocity fields or low-pass
filtering (100 km cutoff) reduces the horizontal divergence by about 70$\%$.
Furthermore, the seasonal cycle in horizontal divergence is out-of-phase with that
of vertical vorticity: the horizontal divergence is about 30$\%$ larger in late
Summer/early Fall than it is in late Winter/early Spring. The divergence associated
with submesoscale subinertial eddies is small and its seasonal cycle is in phase
with that of vorticity (see green lines).  The seasonal cycle of the horizontal
divergence is roughly in phase with the seasonal cycle of the near-surface (0-50m)
stratification.
\begin{figure*}[ht]
\begin{center}
\includegraphics[width=1.\textwidth]{figs/fig1.pdf}
 \caption{Snapshots of surface vorticity (a and b) and a transect
          of potential density at 165$^\circ$E (c and d). The snapshots were
          taken at 00:00 UTC.}
\vspace{-1.5cm}
 \label{fig1}
 \end{center}
 \end{figure*}

 \begin{figure*}[ht]
   \begin{center}
     \includegraphics[width=.9\textwidth]{figs/fig2.pdf}
  \caption{Time series of the root-mean-square of vorticity (a),
  horizontal divergence (b), and stratification in the upper 50 m (c).
  \textcolor{red}{May add a time series of baroclinic production $\overline{w'b'}$ in
  the mixed layer. }.}
  \label{fig2}
  \end{center}
\end{figure*}

@@

\section{Projection on time scales}
To better understand the high-frequency/high-wavenumber flows  we calculate frequency
spectra of surface KE (figure \ref{fig4}). To obtain statistically meaningful results,
we average spectral estimates along a section at 165$^\circ$E.  There is also about 50
independent realizations of spectral estimates.

@@

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=.45\textwidth]{figs/fig4.pdf}
 \caption{Surface (horizontal) KE frequency spectra. Solid lines
 are spectra based on the 1/24$^\circ$ fields, dashed line are spectra
 based on 100-km-smoothed fields.}
 \label{fig4}
 \end{center}
\end{figure}
